---
title: "Module2Discussion1"
author: "Joseph Annand"
date: "2023-07-10"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Discussion

R is a relatively slow programming language, even compared to Python, another interpreted language (https://towardsdatascience.com/python-vs-and-r-for-data-science-4a32580846a4). Specifically, loops in R take comparatively long times to execute. While loops are standard practice in other programming languages, R takes a different approach using vectorized functions. Vectorized functions operate on all elements of a vector at once, saving a considerable amount of time compared to looping through the vector and operating on each element individually (https://docs.ycrc.yale.edu/r-novice-gapminder/09-vectorization/#:~:text=Most%20of%20R's%20functions%20are,read%2C%20and%20less%20error%20prone.&text=The%20multiplication%20happened%20to%20each%20element%20of%20the%20vector). Mathematical operators and some base R functions, like paste, are vectorized.

The example below shows the difference between a vectorized function and a loop (https://towardsdatascience.com/make-your-r-code-10x-faster-vectorization-explained-in-3-minutes-9eb4cdd7a49e):

```{r}
library(microbenchmark)

x <- 1:50

# This vectorized expression...
vectorized <- function(x) {
  paste0("file_", x)
}

# Does the same thing as this loop!
loop <- function(x) {
  filenames <- vector()
  for (i in x) {
    filenames[i] <- paste0("file_", i)
  }
}

microbenchmark(vectorized(x), loop(x))
```

The vectorized function takes an average time of 71 microseconds to execute, while the loop function takes an average time of 354 microseconds to execute.

Vectorized operations allows R programmers to save time when applying functions to large amounts of data. For example, vectorized functions would be useful when it is necessary to perform an operation on all observations in a column of a dataframe. Vectorized functions are also useful for manipulating matrices. In cases where a task is repeatedly performed in a script, it may be beneficial to use a vectorized operation to reduce the execution time of the script.

Scalar functions can be vectorized using the base R function, Vectorize (https://rdrr.io/r/base/Vectorize.html). Implementing this function can be useful when we want to apply some function that typically takes a scalar value as its argument to a large amount of data as a vector. In one example in the R documentation for base::Vectorize, base::rep.int is vectorized. The output of the function changes from a single integer vector to a list of integer vectors, demonstrating another way vectorization can change operations. In another example, Vectorize is applied to a custom function that performs a couple lines of arithmetic. The function is used in building a nonlinear regression contour plot. By vectorizing functions, we can more easily accomplish larger or more advanced tasks.

